title: "Exercise 1"
output: html_document
date: "2024-11-01"
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```
This code chunk will create a function to transform datasets. Basically, the function "transformed_data" will add a column to tha dataset with the information of the row_names, so we can work on this information. Additionally, it will select the new column with the row names, and the first two columns of the dataset.

```{r}
#Loading libraries
suppressPackageStartupMessages({ #Suppressing starting package messages
  library(tidyverse)
  library(dplyr)
})

#Create function "transformed_data" to transform datasets
transformed_data <- function(data) {
  data_2 <- data %>%
    mutate(row_names= rownames(data)) %>% #Add a new column with the names of the rows in order to be able to work with this information
    select(row_names,1:2) #Select only the new column and the 2 first columns of the dataset
  return(data_2)
}
```
Now we can apply the function to the mtcars dataset, which contains information on fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

```{r}
#Loading R-built in dataset "mtcars"
data("mtcars")

# Calling the function and assigning the result to a new dataset "transformed_mtcars"
transformed_mtcars <- transformed_data(mtcars)

# Printing the results
print(transformed_mtcars)
```